---
title: "differential-equations"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

Following the tutorial from:

Differential Equations in R Part 1: Representing Basic Dynamics

https://www.youtube.com/watch?v=1iNXQypailI

Differential Equations in R Part 2: Solving Lotka-Volterra Predation Equations

https://www.youtube.com/watch?v=lJqiasw7OPs

Note: I fixed a few mistakes with the functions where it was using more global scope than I wanted.

Install the library [deSolve](http://desolve.r-forge.r-project.org/) from Packages -> Install

Use the library deSolve and lattice

```{r}
library(deSolve)
library(lattice)
```

Solving the continous equation
$$
\frac{dN}{dt} = rN
$$

Create a function 
```{r}
cgrowth <- function(times, y, parms) {
  r <- parms[1]
  N <- y[1]
  dN.dt <- r * N
  return(list(dN.dt))
}

p  <- c(p=0.5)
y0 <- c(N=2)
t  <- 0:20

sol <- ode(y = y0, times = t, func = cgrowth, parms = p)
sol
```

Plotting it would be
```{r}
plot(sol, type='o', xlab="Time", ylab="Size", main="Solution",
     pch=16, cex=0.7, fg="grey70", col="steelblue2", col.axis="grey30",
     col.lab="grey30", col.main="grey30")
```

Solving a continous time logistic equation numerically

$$
\frac{dN}{dt} = rN(1 - \frac{N}{K})
$$

```{r}
clogistic <- function(times, y, parms) {
  r <- parms[1]
  K <- parms[2]
  N <- y[1]
  dN.dt <- r * N * (1 - (N / K))
  return(list(dN.dt))
}

p  <- c(r=5, K=1000)
y0 <- c(N=2)
t  <- 0:30

sol2 <- ode(y = y0, times = t, func = clogistic, parms = p)
sol2
```

Plotting it would be
```{r}
plot(sol2, type='o', xlab="Time", ylab="Size", main="Solution",
     pch=16, cex=0.7, fg="grey70", col="steelblue2", col.axis="grey30",
     col.lab="grey30", col.main="grey30")
```

Lotka-Volterra Predation Equations
(Coupled Differential Equations)

$$
\frac{dN}{dt} = rN - a PN
$$
$$
\frac{dP}{dt} = -bP + fPN
$$

I use the `with` and `as.list` as short hands to not have to unbox/deconstruct variables
```{r}
predpreyLV <- function(t, y, p) {
 with(c(as.list(y), as.list(p)), {
   dNdt <- r * N - a * P * N
   dPdt <- -b * P + f * P * N
   return(list(c(dNdt, dPdt)))
 })
}

r <- 0.50
a <- 0.01
f <- 0.01
b <- 0.20

p      <- c(r=r, a=a, b=b, f=f)
y0     <- c(N=25, P=5)
times  <- seq(0, 200, 0.1)
LV.out <- ode(y = y0, times, predpreyLV, p)

matplot(LV.out[,1], (LV.out[,2:3]), type="l", xlab="Time", 
        ylab="Population Size",fg="grey70", col=c("darkblue", "red"),
        col.axis="grey30", col.lab="grey30", col.main="grey30")
```

In phase space it would be
```{r}
plot(LV.out[,2], LV.out[,3], type="l", xlab="Prey Density",
     ylab="Predator Density", fg="grey70", col="steelblue2",
     col.axis="grey30", col.lab="grey30", col.main="grey30")
```

Making it more realistic by adding a carryig capacity to prey density

$$
\frac{dN}{dt} = rN[1 - (\frac{N}{K})] - aPN
$$
$$
\frac{dP}{dt} = -bP + fPN
$$

Numerically in R that would be

```{r}
predpreyCarryingLV <- function(t, y, p) {
 with(c(as.list(y), as.list(p)), {
   dNdt <- r * N * (1 - (N / K)) - a * P * N
   dPdt <- -b * P + f * P * N
   return(list(c(dNdt, dPdt)))
 })
}

r <- 0.50
a <- 0.01
f <- 0.01
b <- 0.20
K <- 1000

p      <- c(r=r, a=a, b=b, f=f, K=K)
y0     <- c(N=25, P=5)
times  <- seq(0, 1000, 0.1)
LV.out <- ode(y = y0, times, predpreyCarryingLV, p)

matplot(LV.out[,1], (LV.out[,2:3]), type="l", xlab="Time", 
        ylab="Population Size",fg="grey70", col=c("darkblue", "red"),
        col.axis="grey30", col.lab="grey30", col.main="grey30")
```

As a phase diagram it would be

```{r}
plot(LV.out[,2], LV.out[,3], type="l", xlab="Prey Density",
     ylab="Predator Density", fg="grey70", col="steelblue2",
     col.axis="grey30", col.lab="grey30", col.main="grey30")
```