---
title: "differential-equations"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

Following the tutorial from:

Differential Equations in R Part : Representing Basic Dynamics
https://www.youtube.com/watch?v=1iNXQypailI

Note: I fixed a few mistakes with the functions where it was using more global scope than I wanted.

Install the library [deSolve](http://desolve.r-forge.r-project.org/) from Packages -> Install

Use the library deSolve

```{r}
library(deSolve)
```

Solving the continous equation
$$
\frac{dN}{dt} = rN
$$

Create a function 
```{r}
cgrowth <- function(times, y, parms) {
  r <- parms[1]
  N <- y[1]
  dN.dt <- r * N
  return(list(dN.dt))
}

p  <- 0.5
y0 <- 2
t  <- 0:20

sol <- ode(y = y0, times = t, func = cgrowth, parms = p)
sol
```

Plotting it would be
```{r}
plot(sol, type='o', xlab="Time", ylab="Size", main="Solution",
     pch=16, cex=0.7, fg="grey70", col="steelblue2", col.axis="grey30",
     col.lab="grey30", col.main="grey30")
```

Solving a continous time logistic equation numerically

$$
\frac{dN}{dt} = rN(1 - \frac{N}{K})
$$

```{r}
clogistic <- function(times, y, parms) {
  r <- parms[1]
  K <- parms[2]
  N <- y[1]
  dN.dt <- r * N * (1 - (N/K))
  return(list(dN.dt))
}

p  <- c(5, 1000)
y0 <- 2
t  <- 0:30

sol2 <- ode(y = y0, times = t, func = clogistic, parms = p)
sol2
```

Plotting it would be
```{r}
plot(sol2, type='o', xlab="Time", ylab="Size", main="Solution",
     pch=16, cex=0.7, fg="grey70", col="steelblue2", col.axis="grey30",
     col.lab="grey30", col.main="grey30")
```






